---
title: "Zegveld Bayesian inference result analysis"
author: "Alexander Buzacott"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tibble)
library(readr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(stringr)
library(randomForest)
library(data.table)
library(BayesianTools)
library(future.apply)
library(DescTools)

theme_set(theme_bw())

annotation_custom <- function(grob,
                              xmin = -Inf,
                              xmax = Inf,
                              ymin = -Inf,
                              ymax = Inf,
                              data) {
  layer(data = data, stat = StatIdentity, position = PositionIdentity,
        geom = ggplot2:::GeomCustomAnn,
        inherit.aes = TRUE, params = list(grob = grob,
                                          xmin = xmin, xmax = xmax,
                                          ymin = ymin, ymax = ymax))
}

source("src/modelling_functions.R")
```

This notebook analyses the results of `zegveld_bayesian_fluxes.R` and generates
the summary plots and statistics.

## Load data

```{r}
df <- read_rds(file.path("output", "df.rds"))
```

## Years modelled

```{r}
years <- 2021:2022
```

## CO2 results

### CO2 data

```{r}
co2_model_data <- read_rds(file.path("output", "co2_model_data.rds"))
```

### Read in results

Read in posterior distributions for each year

```{r}
co2_post_dist <- vector("list", length(years))
for (i in seq_along(years)) {
  co2_r_fname <- file.path("output", paste0("co2_r", years[i], ".rds"))
  co2_post_dist[[i]] <- read_rds(co2_r_fname)
}
names(co2_post_dist) <- years
```

### Plot posterior

```{r}
co2_post_dist_p_fname <- file.path("output", "co2_post_dist_p.rds")

if (!file.exists(co2_post_dist_p_fname)) {
  co2_post_dist_p <- lapply(co2_post_dist, function(x) {
    as_tibble(getSample(x))
  })
  co2_post_dist_p <- bind_rows(co2_post_dist_p, .id = "year") %>%
    select(-contains("nee")) %>%
    pivot_longer(-year) %>%
    mutate(type = stringr::str_split_fixed(name, "_", 2)[, 1],
           type = factor(type,
                         levels = c("pt", "rf"),
                         labels = c("Typha", "Ref")),
           param = stringr::str_split_fixed(name, "_", 2)[, 2],
           src = if_else(param %in% c("e_0", "r_ref"), "Reco", "GPP"),
           param = factor(param,
                          levels = c("e_0",
                                     "r_ref",
                                     "alpha",
                                     "beta",
                                     "alpha_amplitude",
                                     "beta_amplitude",
                                     "phase_shift"),
                          labels = c("E[0]",
                                     "R[Ref]",
                                     "alpha",
                                     "beta",
                                     "A[alpha]",
                                     "A[beta]",
                                     "phi")))
  write_rds(co2_post_dist_p, co2_post_dist_p_fname, compress = "gz")
} else {
  co2_post_dist_p <- read_rds(co2_post_dist_p_fname)
}
```

```{r}
p <- co2_post_dist_p %>%
  ggplot(aes(value, col = type, linetype = year)) +
  geom_density() +
  scale_linetype_manual(values = c("2020" = 1, "2021" = 2, "2022" = 3)) +
  facet_wrap(~src + param,
             scales = "free",
             labeller = label_parsed,
             ncol = 5,
             nrow = 2) +
  labs(x = NULL,
       y = "Density",
       col = "Source",
       linetype = "Year") +
  scale_colour_manual(values = c("Typha" = scales::hue_pal()(3)[1],
                                 "Ref" = scales::hue_pal()(3)[2]))

# To be added later
inset <- co2_post_dist_p %>%
  filter(param == "A[beta]") %>%
  ggplot(aes(value, col = type, linetype = year)) +
  geom_density() +
  coord_cartesian(ylim = c(0, 0.5)) +
  guides(col = "none", linetype = "none") +
  labs(x = NULL, y = NULL) +
  theme(plot.background = element_rect(colour = "black", linewidth = 0.5)) +
  scale_colour_manual(values = c("Typha" = scales::hue_pal()(3)[1],
                                 "Ref" = scales::hue_pal()(3)[2])) +
  scale_linetype_manual(values = c("2020" = 1, "2021" = 2, "2022" = 3))

p <- p +
  annotation_custom(ggplotGrob(inset),
                    data = co2_post_dist_p %>% filter(param == "A[beta]"),
                    xmin = 2.5, xmax = 29,
                    ymin = 0.75, ymax = 9)

ggsave(file.path("figures", "co2_post_dist.png"),
       p,
       dpi = 300,
       width = 15,
       height = 6)
```

### Simulate timeseries

```{r}
# Full timeseries with no missing data
co2_eval_data <- df %>%
  select(datetime,
         year,
         doy,
         TS_PT,
         TS_RF,
         Rg = PPFD_IN_F,
         NEE,
         Contribution_GF) %>%
  as.data.table()

# Function to evaluate parameter sets
co2_model_ts <- function(params, data) {
  # Subset data
  data <- data[year == params[1]]

  pt_r_ref <- params[2]
  pt_e_0 <- params[3]
  pt_alpha <- params[4]
  pt_beta <- params[5]
  pt_alpha_amplitude <- params[6]
  pt_beta_amplitude <- params[7]
  pt_phase_shift <- params[8]
  rf_r_ref <- params[9]
  rf_e_0 <- params[10]
  rf_alpha <- params[11]
  rf_beta <- params[12]
  rf_alpha_amplitude <- params[13]
  rf_beta_amplitude <- params[14]
  rf_phase_shift <- params[15]

  pt_reco <- respiration_lloyd_taylor(pt_r_ref, pt_e_0, data$TS_PT)
  pt_gpp <- gpp_sin(
    alpha = pt_alpha,
    beta = pt_beta,
    Rg = data$Rg,
    doy = data$doy,
    phase_shift = pt_phase_shift,
    alpha_amplitude = pt_alpha_amplitude,
    beta_amplitude = pt_beta_amplitude
  )
  pt_nee <- pt_gpp + pt_reco

  rf_reco <- respiration_lloyd_taylor(rf_r_ref, rf_e_0, data$TS_RF)
  rf_gpp <- gpp_sin(
    alpha = rf_alpha,
    beta = rf_beta,
    Rg = data$Rg,
    doy = data$doy,
    phase_shift = rf_phase_shift,
    alpha_amplitude = rf_alpha_amplitude,
    beta_amplitude = rf_beta_amplitude
  )
  rf_nee <- rf_gpp + rf_reco

  nee_sim <- pt_nee * data$Contribution_GF + rf_nee * (1 - data$Contribution_GF)

  dt <- data.table(datetime = data$datetime,
                   pt_reco = pt_reco,
                   pt_gpp = pt_gpp,
                   pt_nee = pt_nee,
                   rf_reco = rf_reco,
                   rf_gpp = rf_gpp,
                   rf_nee = rf_nee,
                   NEE_sim = nee_sim)
  return(dt)
}
```

### Model performance statistics

```{r}
co2_params_full <- lapply(co2_post_dist, function(x) {
  as.data.table(getSample(x))
}) %>%
  bind_rows(.id = "year") %>%
  mutate(year = as.numeric(year))

# Function to return stats from each simulation
co2_stats_f <- function(params, data) {
  ts <- co2_model_ts(params, data)

  obs <- data[year == params[[1]]]$NEE
  sim <- ts$NEE_sim

  perf <- c(
    year = params[[1]],
    RMSE = RMSE(obs, sim),
    `R$^2$` = cor(obs, sim)^2,
    LCCC = CCC(obs, sim)$rho.c$est,
    Bias = mean(obs - sim)
  )

  return(perf)
}
```

```{r}
co2_sim_stats_fname <- file.path("output", "co2_sim_stats.rds")

if (!file.exists(co2_sim_stats_fname)) {
  plan(multisession, workers = 7)
  co2_sim_stats <- future_apply(co2_params_full,
                                MARGIN = 1,
                                co2_stats_f,
                                data = co2_eval_data %>% filter(is.finite(NEE)))
  plan(sequential)
  write_rds(co2_sim_stats, co2_sim_stats_fname, compress = "gz")
} else {
  co2_sim_stats <- read_rds(co2_sim_stats_fname)
}

co2_sim_stats <- as_tibble(t(co2_sim_stats))
```

Write a summary csv

```{r}
co2_sim_stats_summary_fname <- file.path("results", "co2_sim_stats_summary.csv")

if (!file.exists(co2_sim_stats_summary_fname)) {
  co2_sim_stats_summary <- co2_sim_stats %>%
    pivot_longer(c(RMSE, `R$^2$`, LCCC, Bias)) %>%
    group_by(name) %>%
    summarise(min = min(value),
              q02.5 = quantile(value, 0.025),
              q05 = quantile(value, 0.05),
              q25 = quantile(value, 0.25),
              mean = mean(value),
              median = median(value),
              q75 = quantile(value, 0.75),
              q95 = quantile(value, 0.95),
              q97.5 = quantile(value, 0.975),
              max = max(value),
              sd = sd(value)) %>%
    pivot_longer(-name, names_to = "stat") %>%
    pivot_wider()

  write_csv(co2_sim_stats_summary, co2_sim_stats_summary_fname)
} else {
  co2_sim_stats_summary <- read_csv(co2_sim_stats_summary_fname)
}

co2_sim_stats_summary %>%
  mutate(across(where(is.numeric), ~round(.x, 2))) %>%
  knitr::kable()
```

### Model performance objects

```{r}
co2_sim_fname <- file.path("output", "co2_sim.rds")
co2_sim_day_fname <- file.path("output", "co2_sim_day.rds")
co2_obs_sim_fname <- file.path("output", "co2_obs_sim.rds")

if (!file.exists(co2_sim_fname)) {
  # Take sample of 3000 sets for plotting
  co2_params_sample <- lapply(co2_post_dist, function(x) {
    as.data.table(getSample(x, numSamples = 3000))
  })
  co2_params_sample <- bind_rows(co2_params_sample, .id = "year") %>%
    mutate(year = as.numeric(year))

  plan(multisession, workers = 7)
  co2_sim <- future_apply(co2_params_sample,
                          MARGIN = 1,
                          co2_model_ts,
                          data = co2_eval_data)
  plan(sequential)

  write_rds(co2_sim, co2_sim_fname, compress = "gz")
} else {
  co2_sim <- read_rds(co2_sim_fname)
}

if (!file.exists(co2_sim_day_fname)) {
  co2_sim_day <- lapply(co2_sim, function(x) {
    x$datetime <- as_date(x$datetime - 1)
    x <- x[, lapply(.SD, mean), by = datetime]
    x[, 2:ncol(x)] <- x[, 2:ncol(x)] * 44 * 10^-9 * 10^4 * 86400 # µmol/m2/s -> t CO2/ha/day
    return(x)
  }) %>%
    bind_rows(.id = "Sim")
  write_rds(co2_sim_day, co2_sim_day_fname, compress = "gz")
} else {
  co2_sim_day <- read_rds(co2_sim_day_fname)
}

if (!file.exists(co2_obs_sim_fname)) {
  co2_obs_sim_summary <- rbindlist(co2_sim) %>%
    filter(datetime %in% co2_model_data$datetime) %>%
    group_by(datetime) %>%
    summarise(
    min = min(NEE_sim),
    q02.5 = quantile(NEE_sim, 0.025),
    q25 = quantile(NEE_sim, 0.25),
    mean = mean(NEE_sim),
    median = median(NEE_sim),
    q75 = quantile(NEE_sim, 0.75),
    q97.5 = quantile(NEE_sim, 0.975),
    max = max(NEE_sim)
  ) %>%
  left_join(df %>% select(datetime, night, Obs = NEE))
  write_rds(co2_obs_sim_summary, co2_obs_sim_fname, compress = "gz")
} else {
  co2_obs_sim_summary <- read_rds(co2_obs_sim_fname)
}
```

### Plot of performance

```{r}
p1 <- ggplot(co2_obs_sim_summary) +
  geom_point(aes(Obs, mean), size = 0.5) +
  geom_abline(col = "red") +
  labs(x = expression(Obs~NEE~"["*µmol~m^{-2}~s^{-1}*"]"),
       y = expression(Sim~NEE~"["*µmol~m^{-2}~s^{-1}*"]")) +
  scale_y_continuous(limits = c(-30, 30),
                     breaks = seq(-30, 30, 10)) +
  scale_x_continuous(limits = c(-30, 30),
                     breaks = seq(-30, 30, 10))

p2 <- co2_obs_sim_summary %>%
  filter(night) %>%
  ggplot() +
  geom_point(aes(Obs, mean), size = 0.5) +
  geom_abline(col = "red") +
  labs(x = expression(Obs~night~NEE~"["*µmol~m^{-2}~s^{-1}*"]"),
       y = expression(Sim~night~NEE~"["*µmol~m^{-2}~s^{-1}*"]")) +
  ylim(0, NA)

p3 <- co2_obs_sim_summary %>%
  filter(!night) %>%
  ggplot() +
  geom_point(aes(Obs, mean), size = 0.5) +
  geom_abline(col = "red") +
  labs(x = expression(Obs~day~NEE~"["*µmol~m^{-2}~s^{-1}*"]"),
       y = expression(Sim~day~NEE~"["*µmol~m^{-2}~s^{-1}*"]")) +
  scale_y_continuous(limits = c(-30, 30),
                     breaks = seq(-30, 30, 10)) +
  scale_x_continuous(limits = c(-30, 30),
                     breaks = seq(-30, 30, 10))
p <- egg::ggarrange(p1, p2, p3, nrow = 1)

ggsave(file.path("figures", "co2_obs_sim.png"),
       p,
       dpi = 300,
       width = 9,
       height = 3)
```

### Model sim timeseries

```{r}
co2_sim_summary_day <- co2_sim_day %>%
  select(Sim, date = datetime, pt_nee, rf_nee, NEE_sim) %>%
  pivot_longer(-c(date, Sim)) %>%
  mutate(name = factor(name,
                       levels = c("pt_nee", "rf_nee", "NEE_sim"),
                       labels = c("Typha", "Ref", "NEE_sim"))) %>%
  group_by(date, name) %>%
  summarise(
    min = mean(value),
    q02.5 = quantile(value, 0.025, na.rm = TRUE),
    q05 = quantile(value, 0.05, na.rm = TRUE),
    q25 = quantile(value, 0.25, na.rm = TRUE),
    mean = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    q75 = quantile(value, 0.75, na.rm = TRUE),
    q95 = quantile(value, 0.95, na.rm = TRUE),
    q97.5 = quantile(value, 0.975, na.rm = TRUE),
    max = mean(value),
    sd = max(value),
    .groups = "drop"
  ) %>%
  mutate(year = year(date),
         doy = yday(date))

write_rds(co2_sim_summary_day,
          file.path("output", "co2_sim_summary_day.rds"), compress = "gz")
```

```{r}
# CO2 daily observed data
co2_obs_day <- df %>%
  mutate(date = as_date(datetime - 1)) %>%
  select(date, night, NEE, NEE_GF = NEE_F_MDS) %>%
  group_by(date) %>%
  summarise(n = sum(!is.na(NEE)),
            n_night = sum(night & !is.na(NEE)),
            n_day = n - n_night,
            n_night_hours = sum(night),
            n_day_hours = sum(!night),
            mean = mean(NEE, na.rm = TRUE),
            mean_day = mean(NEE[!night], na.rm = TRUE),
            mean_night = mean(NEE[night], na.rm = TRUE),
            mean2 = mean_day + mean_night,
            mean_gf = mean(NEE_GF, na.rm = TRUE),
            .groups = "drop") %>%
  # filter(n > 24) %>%
  mutate(across(c(mean, mean_gf, mean2),
                ~.x * 44 * 10^-6 * 10^-3 * 10^4 * 86400)) %>%
  mutate(year = year(date),
         year_gr = year(date %m-% months(2)),
         doy = yday(date))

write_rds(co2_obs_day, file.path("output", "co2_obs_day.rds"), compress = "gz")
```

#### Plot

```{r}
ggplot(co2_sim_summary_day) +
  geom_point(data = co2_obs_day, aes(doy, mean_gf, fill = "Obs"), size = 0.5) +
  geom_ribbon(aes(x = doy, ymin = q02.5, ymax = q97.5, col = name), alpha = 0.5, show.legend = FALSE, linewidth = 0.5) +
  geom_line(aes(doy, median, col = name)) +
  facet_wrap(~year, ncol = 1) +
  labs(x = "Day of the year",
       col = NULL,
       fill = NULL,
       y = expression(NEE~"["*kg~ha^{-1}*~d^{-1}*"]")) +
  scale_colour_discrete(labels = c("Typha", "Ref", expression(NEE~Sim))) +
  scale_fill_manual(values = c("Obs" = 1), label = expression(NEE)) +
  theme(legend.text.align = 0)
```

### Budget

Use all simulations for the budget

```{r}
co2_budget_f <- function(params, data) {
  ts <- co2_model_ts(params, data)
  ys <- 44 * 10^-8 * 1800 # µmol/m2/s -> kg CO2/ha/1800 for summing

  dt <- c(year = params[[1]],
          pt_reco = sum(ts$pt_reco) * ys,
          pt_gpp = sum(ts$pt_gpp) * ys,
          pt_nee = sum(ts$pt_nee) * ys,
          rf_reco = sum(ts$rf_reco) * ys,
          rf_gpp = sum(ts$rf_gpp) * ys,
          rf_nee = sum(ts$rf_nee) * ys,
          NEE_sim = sum(ts$NEE_sim) * ys)

  return(dt)
}
```

```{r}
co2_sim_budget_fname <- file.path("output", "co2_sim_budget.rds")

if (!file.exists(co2_sim_budget_fname)) {
  plan(multisession, workers = 7)
  co2_sim_budget <- future_apply(co2_params_full,
                                 MARGIN = 1,
                                 co2_budget_f,
                                 data = co2_eval_data)
  plan(sequential)

  co2_sim_budget <- as_tibble(t(co2_sim_budget)) %>%
    pivot_longer(-year) %>%
    group_by(year, name) %>%
    summarise(min = min(value),
              q02.5 = quantile(value, 0.025),
              q05 = quantile(value, 0.05),
              q25 = quantile(value, 0.25),
              mean = mean(value),
              sd = sd(value),
              q75 = quantile(value, 0.75),
              q95 = quantile(value, 0.95),
              q97.5 = quantile(value, 0.975),
              max = max(value),
              .groups = "drop") %>%
    pivot_longer(-c(year, name), names_to = "stat") %>%
    pivot_wider()

  write_rds(co2_sim_budget, co2_sim_budget_fname)
} else {
  co2_sim_budget <- read_rds(co2_sim_budget_fname)
}
```

```{r}
co2_sim_budget %>%
  mutate(across(-c(year, stat), ~round(.x, 1))) %>%
  knitr::kable()
```

### Reco/GPP Timeseries

```{r}
co2_fp_summary_day <- co2_sim_day %>%
  select(date = datetime, pt_reco, pt_gpp, rf_gpp, rf_reco) %>%
  pivot_longer(-date) %>%
  group_by(date, name) %>%
  summarise(
    min = mean(value),
    q02.5 = quantile(value, 0.025, na.rm = TRUE),
    q05 = quantile(value, 0.05, na.rm = TRUE),
    q25 = quantile(value, 0.25, na.rm = TRUE),
    mean = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    q75 = quantile(value, 0.75, na.rm = TRUE),
    q95 = quantile(value, 0.95, na.rm = TRUE),
    q97.5 = quantile(value, 0.975, na.rm = TRUE),
    max = mean(value),
    sd = max(value),
    .groups = "drop"
  ) %>%
  mutate(year = year(date),
         doy = yday(date)) %>%
  mutate(src = str_sub(name, 1, 2),
         name = str_sub(name, 4, 7))
```

```{r}
ggplot(co2_fp_summary_day, aes(date, mean, col = src)) +
  geom_ribbon(aes(date, ymin = q02.5, ymax = q97.5)) +
  geom_line() +
  facet_wrap(~name, ncol = 1)
```

## CH4 results

### CH4 data

```{r}
ch4_model_data <- read_rds(file.path("output", "ch4_model_data.rds"))
```

### Read in results

```{r}
ch4_post_dist <- vector("list", length(years))
for (i in seq_along(years)) {
  ch4_r_fname <- file.path("output", paste0("ch4_r", years[i], ".rds"))
  ch4_post_dist[[i]] <- read_rds(ch4_r_fname)
}
names(ch4_post_dist) <- years
```

### Plot posterior

```{r}
ch4_post_dist_p_fname <- file.path("output", "ch4_post_dist_p.rds")
if (!file.exists(ch4_post_dist_p_fname)) {
  ch4_post_dist_p <- lapply(ch4_post_dist, function(x) {
    as_tibble(getSample(x))
  })
  ch4_post_dist_p <- bind_rows(ch4_post_dist_p, .id = "year") %>%
    select(-contains("err")) %>%
    pivot_longer(-year) %>%
    mutate(type = stringr::str_split_fixed(name, "_", 2)[, 1],
           type = factor(type,
                         levels = c("pt", "rf"),
                         labels = c("Typha", "Ref")),
           param = stringr::str_split_fixed(name, "_", 2)[, 2],
           param = factor(param,
                          levels = c("a", "b", "k")))
  write_rds(ch4_post_dist_p, ch4_post_dist_p_fname, compress = "gz")
} else {
  ch4_post_dist_p <- read_rds(ch4_post_dist_p_fname)
}
```

```{r}
p <- ch4_post_dist_p %>%
  drop_na() %>%
  ggplot(aes(value, col = type, linetype = year)) +
  geom_density() +
  scale_linetype_manual(values = c("2021" = 2, "2022" = 3)) +
  facet_wrap(~param,
             scales = "free",
             labeller = label_parsed,
             ncol = 5,
             nrow = 2) +
  labs(x = NULL,
       y = "Density",
       col = "Source",
       linetype = "Year") +
  scale_colour_manual(values = c("Typha" = scales::hue_pal()(3)[1],
                                 "Ref" = scales::hue_pal()(3)[2]))

ggsave(file.path("figures", "ch4_post_dist.png"),
       p,
       dpi = 300,
       width = 10,
       height = 3)
```

### Simulate timeseries

```{r}
ch4_eval_data <- df %>%
  select(datetime,
         year,
         season,
         night,
         doy,
         night,
         Contribution_GF,
         FCH4,
         TS_PT,
         TS_RF,
         WL_PT,
         WL_RF) %>%
  as.data.table()

ch4_model_ts <- function(params, data) {
  data <- data[year == params[1]]
  pt_pred <- fch4_temp_wl(data$TS_PT, data$WL_PT,
                          params[2], params[3], params[4])
  rf_pred <- fch4_temp_wl(data$TS_RF, data$WL_RF,
                          params[5], params[6], params[7])

  nee_ch4_pred <- pt_pred * data$Contribution_GF +
    rf_pred * (1 - data$Contribution_GF)

  dt <- data.table(datetime = data$datetime,
                   pt_fch4 = pt_pred,
                   rf_fch4 = rf_pred,
                   FCH4_sim = nee_ch4_pred)

  return(dt)
}
```

### Model performance statistics

```{r}
ch4_params_full <- lapply(ch4_post_dist, function(x) {
  as.data.table(getSample(x))
}) %>%
  bind_rows(.id = "year") %>%
  mutate(year = as.numeric(year))

# Function to return stats from each simulation
ch4_stats_f <- function(params, data) {
  ts <- ch4_model_ts(params, data)

  obs <- data[year == params[[1]]]$FCH4
  sim <- ts$FCH4_sim

  perf <- c(
    year = params[[1]],
    RMSE = RMSE(obs, sim),
    `R$^2$` = cor(obs, sim)^2,
    LCCC = CCC(obs, sim)$rho.c$est,
    Bias = mean(obs - sim)
  )

  return(perf)
}
```

```{r}
ch4_sim_stats_fname <- file.path("output", "ch4_sim_stats.rds")

if (!file.exists(ch4_sim_stats_fname)) {
  plan(multisession, workers = 7)
  ch4_sim_stats <- future_apply(ch4_params_full,
                                MARGIN = 1,
                                ch4_stats_f,
                                data = filter(ch4_eval_data, is.finite(FCH4)))
  plan(sequential)
  write_rds(ch4_sim_stats, ch4_sim_stats_fname, compress = "gz")
} else {
  ch4_sim_stats <- read_rds(ch4_sim_stats_fname)
}

ch4_sim_stats <- as_tibble(t(ch4_sim_stats))
```

Write a summary csv

```{r}
ch4_sim_stats_summary_fname <- file.path("results", "ch4_sim_stats_summary.csv")

if (!file.exists(ch4_sim_stats_summary_fname)) {
  ch4_sim_stats_summary <- ch4_sim_stats %>%
    pivot_longer(c(RMSE, `R$^2$`, LCCC, Bias)) %>%
    group_by(name) %>%
    summarise(min = min(value),
              q05 = quantile(value, 0.05),
              q02.5 = quantile(value, 0.025),
              q25 = quantile(value, 0.25),
              mean = mean(value),
              median = median(value),
              q75 = quantile(value, 0.75),
              q95 = quantile(value, 0.95),
              q97.5 = quantile(value, 0.975),
              max = max(value),
              sd = sd(value)) %>%
    pivot_longer(-name, names_to = "stat") %>%
    pivot_wider()

  write_csv(ch4_sim_stats_summary, ch4_sim_stats_summary_fname)
} else {
  ch4_sim_stats_summary <- read_csv(ch4_sim_stats_summary_fname)
}

ch4_sim_stats_summary %>%
  mutate(across(where(is.numeric), ~round(.x, 3))) %>%
  knitr::kable()
```

### Model performance objects

```{r}
ch4_sim_fname <- file.path("output", "ch4_sim.rds")
ch4_sim_day_fname <- file.path("output", "ch4_sim_day.rds")
ch4_obs_sim_fname <- file.path("output", "ch4_obs_sim.rds")

if (!file.exists(ch4_sim_day_fname) & !file.exists(ch4_obs_sim_fname)) {
  if (!file.exists(ch4_sim_fname)) {
    ch4_params <- lapply(ch4_post_dist, function(x) {
      as.data.table(getSample(x, numSamples = 3000))
    })
    ch4_params <- bind_rows(ch4_params, .id = "year") %>%
      mutate(year = as.numeric(year))

    ch4_sim <- apply(ch4_params, MARGIN = 1, ch4_model_ts, data = ch4_eval_data)

    write_rds(ch4_sim, ch4_sim_fname, compress = "gz")
  } else {
    ch4_sim <- read_rds(ch4_sim_fname)
  }
}

if (!file.exists(ch4_sim_day_fname)) {
  ch4_sim_day <- lapply(ch4_sim, function(x) {
    x$datetime <- as_date(x$datetime - 1)
    x <- x[, lapply(.SD, mean), by = datetime]
    x[, 2:ncol(x)] <- x[, 2:ncol(x)] * 16 * 10^-9 * 10^4 * 86400 # umol/m2/s -> kg/ha/day
    return(x)
  }) %>%
    bind_rows(.id = "Sim")
  write_rds(ch4_sim_day, ch4_sim_day_fname, compress = "gz")
} else {
  ch4_sim_day <- read_rds(ch4_sim_day_fname)
}

if (!file.exists(ch4_obs_sim_fname)) {
  ch4_obs_sim_summary <- rbindlist(ch4_sim) %>%
    filter(datetime %in% ch4_model_data$datetime) %>%
    group_by(datetime) %>%
    summarise(
      min = min(FCH4_sim),
      q02.5 = quantile(FCH4_sim, 0.025),
      q25 = quantile(FCH4_sim, 0.25),
      mean = mean(FCH4_sim),
      median = median(FCH4_sim),
      q75 = quantile(FCH4_sim, 0.75),
      q97.5 = quantile(FCH4_sim, 0.975),
      max = max(FCH4_sim),
      sd = sd(FCH4_sim)
    ) %>%
    left_join(df %>% select(datetime, night, Obs = FCH4))
  write_rds(ch4_obs_sim_summary, ch4_obs_sim_fname, compress = "gz")
} else {
  ch4_obs_sim_summary <- read_rds(ch4_obs_sim_fname)
}
```

### Plot of performance

```{r}
p1 <- ggplot(ch4_obs_sim_summary) +
  geom_point(aes(Obs, mean), size = 0.5) +
  geom_abline(col = "red") +
  labs(x = expression(Obs~FCH[4]~"["*µmol~m^{-2}~s^{-1}*"]"),
       y = expression(Sim~FCH[4]~"["*µmol~m^{-2}~s^{-1}*"]")) +
  ylim(0, 0.5) +
  xlim(0, 0.5)

ggsave(file.path("figures", "ch4_obs_sim.png"),
       p1,
       dpi = 300,
       width = 3,
       height = 3)
```

### Model sim timeseries

```{r}
ch4_sim_summary_day <- ch4_sim_day %>%
  select(Sim, date = datetime, pt_fch4, rf_fch4, FCH4_sim) %>%
  pivot_longer(-c(date, Sim)) %>%
  mutate(name = factor(name,
                       levels = c("pt_fch4", "rf_fch4", "FCH4_sim"),
                       labels = c("Typha", "Ref", "FCH4_sim"))) %>%
  group_by(date, name) %>%
  summarise(n = n(),
            min = min(value, na.rm = TRUE),
            q01 = quantile(value, 0.01, na.rm = TRUE),
            q02.5 = quantile(value, 0.025, na.rm = TRUE),
            q05 = quantile(value, 0.05, na.rm = TRUE),
            q25 = quantile(value, 0.25, na.rm = TRUE),
            mean = mean(value, na.rm = TRUE),
            median = median(value, na.rm = TRUE),
            q75 = quantile(value, 0.75, na.rm = TRUE),
            q95 = quantile(value, 0.95, na.rm = TRUE),
            q97.5 = quantile(value, 0.975, na.rm = TRUE),
            q99 = quantile(value, 0.99, na.rm = TRUE),
            max = max(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE),
            .groups = "drop") %>%
  mutate(year = year(date),
         doy = yday(date))

write_rds(ch4_sim_summary_day,
          file.path("output", "ch4_sim_summary_day.rds"),
          compress = "gz")
```

```{r}
# Daily aggregate of ch4 observations
ch4_obs_day <- df %>%
  mutate(date = as_date(datetime - 1)) %>%
  select(date, FCH4, FCH4_F_MDS) %>%
  drop_na() %>%
  group_by(date) %>%
  summarise(min = min(FCH4),
            sd = sd(FCH4),
            mean = mean(FCH4),
            median = median(FCH4),
            max = max(FCH4),
            mean_mds = mean(FCH4_F_MDS),
            .groups = "drop") %>%
  mutate(across(where(is.numeric),
                ~.x * 16 * 10^-6 * 10^-3 * 10^4 * 86400)) %>% # µmol/m2/s to t CO2-eq/ha/day
  mutate(year = year(date),
         year_gr = year(date %m-% months(2)),
         doy = yday(date))

write_rds(ch4_obs_day,
          file.path("output", "ch4_obs_day.rds"),
          compress = "gz")
```

#### Plot

```{r}
ch4_sim_summary_day %>%
  ggplot() +
  geom_point(data = ch4_obs_day, aes(doy, mean_mds, fill = "Obs"),
             size = 0.5) +
  geom_ribbon(aes(x = doy, ymin = q02.5, ymax = q97.5, col = name),
              alpha = 0.5,
              show.legend = FALSE,
              linewidth = 0.5) +
  geom_line(aes(doy, median, col = name)) +
  facet_wrap(~year, ncol = 1) +
  labs(x = "Day of the year",
       col = NULL,
       fill = NULL,
       y = expression(FCH[4]~"["*kg~ha^{-1}*~d^{-1}*"]")) +
  # guides(fill = FALSE) +
  scale_colour_discrete(labels = c("Typha", "Ref", expression(FCH[4]~Sim))) +
  scale_fill_manual(values = c("Obs" = 1), label = expression(FCH[4])) +
  theme(legend.text.align = 0) +
  ylim(-0.5, 6)

ggsave(file.path("figures", "ch4_modelled_ts.png"),
       dpi = 300,
       height = 6,
       width = 8)
```

### Budget

```{r}
ch4_budget_f <- function(params, data) {

  ts <- ch4_model_ts(params, data)
  ys <- 16 * 10^-5 * 1800 # µmol/m2/s -> kg CH4/ha/30 min for summing

  dt <- c(year = params[[1]],
          pt_fch4 = sum(ts$pt_fch4) * ys,
          rf_fch4 = sum(ts$rf_fch4) * ys,
          FCH4_sim = sum(ts$FCH4_sim) * ys)

  return(dt)
}
```

```{r}
ch4_sim_budget_fname <- file.path("output", "ch4_sim_budget.rds")

if (!file.exists(ch4_sim_budget_fname)) {
  ch4_params_full <- lapply(ch4_post_dist, function(x) {
    as.data.table(getSample(x))
  }) %>%
    bind_rows(.id = "year") %>%
    mutate(year = as.numeric(year))

  plan(multisession, workers = 7)
  ch4_sim_budget <- future_apply(ch4_params_full,
                                 MARGIN = 1,
                                 ch4_budget_f,
                                 data = ch4_eval_data)
  plan(sequential)

  ch4_sim_budget <- as_tibble(t(ch4_sim_budget)) %>%
    pivot_longer(-year) %>%
    group_by(year, name) %>%
    summarise(
      min = min(value),
      q02.5 = quantile(value, 0.025),
      q05 = quantile(value, 0.05),
      q25 = quantile(value, 0.25),
      mean = mean(value),
      sd = sd(value),
      q75 = quantile(value, 0.75),
      q95 = quantile(value, 0.95),
      q97.5 = quantile(value, 0.975),
      max = max(value),
      .groups = "drop"
    ) %>%
    pivot_longer(-c(year, name), names_to = "stat") %>%
    pivot_wider()

  write_rds(ch4_sim_budget, ch4_sim_budget_fname)
} else {
  ch4_sim_budget <- read_rds(ch4_sim_budget_fname)
}
```

```{r}
ch4_sim_budget %>%
  mutate(across(-c(year, stat), ~round(.x, 1))) %>%
  knitr::kable()
```

```{r}
ch4_sim_budget %>%
  mutate(across(-c(year, stat), ~round(.x * 27 / 1000, 1))) %>%
  knitr::kable()
```
